The documentation (here)[https://cran.r-project.org/web/packages/nametagger/nametagger.pdf] is here.

some hypothesis about the modified sentence `"Hello I am Marianna and this is Mary Jane who lives in Scotland. We work as teachers. We have a cat George and she is a tabby. Scotland is cold and rainy in the winter like the rest of United Kingdom, but I like Scotland."`

First let's see this new sentence, with the old code (slightly improved)

```{r}
#we create a function to run the entity tagger
extract_entities <- function(text) {
  result <- predict(nametagger_model, text, split = '\\s+')  
  result
}

# for my own use, here I create a better data format so the changes are easier to see:
show_only_recognised <- function(text){
  extract_entities(text) %>% # notice we are using one of our functions inside of another of our functions! that's common in programming.
    filter(entity != 'O') %>%
    mutate(full_name = paste(term, entity, sep = "/")) %>%
    summarize(all_names = paste(full_name, collapse = ", "))  # Combine into a single string
}

text ="Hello I am Marianna and this is Mary Jane who lives in Scotland. We work as teachers. We have a cat George and she is a tabby. Scotland is cold and rainy in the winter like the rest of United Kingdom, but I like Scotland."
show_only_recognised(text)
```
Hypothesis (some of which we can fix or test, some we can't without re-training the model):
- 'Scotland.' (with a full stop) is not recognised a s acountry. We could remove pubcluation.
- 'Mary' is not recognised as a name, maybe because tokens are lower-cased, and 'mary' means 'happy'? To fix that we might need to re-train the model to not ignore capitalisation?
- 'Meilo' is an uncommon name
- 'United Kingdom' because it has two words?
- 'Mary Jane' - not su sure? all of the above?

Since code is not a part of this activity, we won't be fixing those. but have a look at this: if all we change is how we separate tokens, using `\\W` (non-word characters) rather than `\\s+` (spaces)... we already fix the problem with the dot in `Scotland.` (last item).


```{r}

text ="Hello I am Mary and this is Paul who lives in Scotland. We work as teachers.  We have a cat Meilo and she is a tabby. Scotland is cold and rainy in the winter but I like Scotland."
predict(nametagger_model, text, split = '\\W')  
```


